CRUSE to Safe Cycling in Ireland \\\large{An Open Source Tool to Support Active Travel}

### 2023-10-10 paper.tex
Incorporate changes from @eugenividal

\begin{abstract}
Legislation mandates improved road safety and decarbonisation of
predominantly road-based transport network in Ireland and other
countries. The Cycle Route Uptake and Scenario Estimation (CRUSE) Tool
was commissioned to support cycle network planning, safety
interventions, business case development and project evaluation. Cycling
in Ireland represents only 3\% of total modal share as of the latest
Census (2016), but accounts for 20\% of serious injuries and 7\% of all
fatalities. High resolution datasets on baseline and potential cycling
levels are needed to meet Ireland's climate and safety targets.

The CRUSE Tool is based on open source software and reproducible data
science. The resulting web application is available at
\url{https://cruse.bike/}, enabling planners, engineers, and other
stakeholders to make more evidence-based decisions. CRUSE goes beyond
previous work in several important ways. It models networks at high
levels of geographic resolution, simulates a wide range of trips
including recreational, social, shopping and personal trips, in addition
to trips to work and primary, secondary and tertiary education
locations. The tool provides key `denominator' information to enable
reporting of collision rates per billion km cycled.

@@

\begin{keyword}
    Cycling \sep Open-source \sep Road Safety \sep 
    Active Travel
\end{keyword}
\end{frontmatter}
    \ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[borderline west={3pt}{0pt}{shadecolor}, interior hidden, boxrule=0pt, frame hidden, breakable, enhanced, sharp corners]}{\end{tcolorbox}}\fi

@@

Transport is responsible for 23\% of global greenhouse gas (GHG)
emissions, 70\% of which are from road transport. Nearly nearly half of
these emissions (around 10\% of global emissions) are from passenger
cars \citep{jaramillo2022}. The transport system encourages, enables and
in some cases enforces unsustainable lifestyles. Services that are only
accessible by car lock-in car dependency
\citep{gray2001, shergold2012, motte-baumvol2010}.

@@

Regarding road traffic casualties, another deadly consequence of
inefficient transport systems, the Road Infrastructure Safety Management
(RISM) directive (2008/96/EC) requires member states to implement a road
safety management system (RSMS) for all public roads. Specifically,
``Member States shall ensure that the ranking of high accident
concentration sections and the network safety ranking are carried out''
\citep{directiv2008}. Given that `safety' in this context is usefully
quantified as the number of people killed and seriously injured (KSI)
per distance travelled, the directive requires estimation of distance
travelled by mode, down to the road link level. For active modes, about
which there is a paucity of data compared with motorized modes, this is
a major challenge. Better data to inform road safety policies and
interventions is a motivation for the CRUSE Tool outlined in this paper,
and modeling active travel more broadly.

@@

At the regional level, the recently published Greater Dublin Area
Transport Strategy \citep{greater} reinforces these findings: ``nearly a
quarter of adults cycle at least once a week in the Dublin Metropolitan
Area'' with cycling in the Dublin area taking up to 60,000 cars off the
road today. Extrapolating this on a per population basis across Ireland,
with around 40\% of the population living in Dublin, this suggests that
around 150,000 cars could be removed nationwide just by achieving Dublin
levels of cycling in all counties (notwithstanding existing cycling
trips and differences in trip distances). 7 in 10 trips in Ireland are
by car \citep{national} . Cycling has the potential to replace a large
proportion of these trips: ``A high priority must also be given to
cyclists, because trips by this mode have the potential to replace trips
by private car, most specifically for short to medium distance trips,
but increasingly for longer trips as e-bikes extend the range of this
mode'' \citep{greater}.

@@

This paper aims to describe the design, features and potential use of
the CRUSE Tool. In Section~\ref{sec-methods}, we outline the methods
used to generate the evidence presented in the CRUSE Tool for Ireland.
In Section~\ref{sec-results}, we present the results of the CRUSE Tool,
including estimates of current cycling levels and future cycling
potential at the national, regional and local levels. In
Section~\ref{sec-discussion}, we discuss the implications of the results
for policy and practice. In Section~\ref{sec-conclusions}, we conclude
the paper.

@@

The methods used to generate the evidence presented in the CRUSE Tool
for Ireland build on the Propensity to Cycle Tool (PCT), which was
originally funded by the UK's Department for Transport and developed by
a multi-university team. An important feature of the PCT is that it is
open source and publicly available (at
\href{https://www.pct.bike/}{www.pct.bike}), allowing its use by all
stakeholders in the transport planning process \citep{lovelace2017}. The
PCT approach has had major policy impacts, as outlined in Research
Excellence Framework (REF) impact case studies, which demonstrate that
the tool ``revolutionised strategic cycle planning in England and
Wales'' by overcoming the barriers to cycling investment imposed by lack
of evidence on cycling potential\citep{lovelace2023}.

The first version of the PCT was based on current and future potential
uptake of cycling for single stage \emph{travel to work} at desire line,
zone, route, and route network levels \citep{lovelace2016}. It was
launched in April 2017 as the government-endorsed tool for strategic
cycle network planning and part of the Cycling and Walking Investment
Strategy \citep{cycling2017}. Extensions of the PCT approach have
included estimation of benefits at the individual level
\citep{woodcock2018}, addition of travel to school network
\citep{goodman2019}, and improved modelling of impacts on health,
environmental and distributional outcomes \citep{woodcock2021}.
Initially developed just for England, the PCT was extended to cover all
of Wales (for commuter data only) in 2018.

@@

A feature of active travel interventions is that they require dense
networks of routes to be effective \citep{parkin2018}. This means that
high levels of geographic resolution are needed in the data used to
estimate cycling potential. However, datasets on travel patterns are
often only available at the level of administrative zones. The Central
Statistics Office (CSO) in Ireland provides Place of Work, School or
College Census of Anonymised Records
(\href{https://www.cso.ie/en/census/census2016reports/powscar/}{POWSCAR})
data on the number of people travelling to work and school at the
Electoral Division (ED) level, for example.

The method used to convert OD data to route networks used in the PCT was
to calculate a single route between the population weighted centroids
associated with each OD pair. This method works fine when the OD data
represents movement between small areas, but was not appropriate for
generating route networks from the POWSCAR data because zone centroids
are so far apart that the resulting route networks would be sparse and
unrealistic. To tackle this issue we developed a new method for OD data
disaggregation called `jittering' \citep{lovelace2022}. The method works
by first disaggregating the OD data based on a `disaggregation
threshold' and then assigning each disaggregated `sub-OD' pair to
`subpoints' within each zone. As shown in Figure~\ref{fig-dublin}, the
resulting route networks are dense, even in rural areas.

@@

However, travel to school and work only constitute around only 30\% of
all trips in Ireland. To overcome this issue, we developed a spatial
interaction modelling methodology to estimate the number of trips
between each OD pair for additional trip purposes. The classification of
trip purposes used in the CRUSE Tool was guided mainly by the trip
purpose classification found within the National Household Travel Survey
(NHTS), but with the addition of categories based on the comprehensive
POWSCAR data, and the need to include recreational trips and multi-stage
trips (not yet implemented). An overview of the trip purposes used in
CRUSE is presented in Figure~\ref{fig-trip-purposes}.

@@

We calculated three network types for each disaggregated (`jittered') OD
pair: `Fastest', `Balanced' and `Quietest'. As outlined on the
CycleStreets'
\href{https://www.cyclestreets.net/help/journey/howitworks/}{website},
the `Fastest' routes minimise journey time, accounting for traffic
lights and surface type. The `Quietest' route type minimises busy
sections of road, allowing high `diversion factors' from the fastest
route on quieter but less direct ways (often avoiding roads and
interactions with motor vehicles altogether where possible). The
`Balanced' route type is a compromise between the two, minimising
journey time while avoiding the busiest roads. Allowing users to switch
between these three route types enables them to consider the trade-offs
between directness and `cycle friendliness' when planning new
infrastructure, as shown in Figure~\ref{fig-route-types}.

@@

The datasets outlined above were used to generate estimates of cycling
currently (the `Baseline' scenario) and under four scenarios of cycling
potential: Near Market, Climate Action Plan, Go Dutch, and Ebike. Each
of these scenarios is described below, as illustrated in Figure
Figure~\ref{fig-scenarios}.

@@

The Baseline scenario approximates current cycling levels. As outlined
in Section~\ref{sec-trip-purposes}, the Baseline scenario includes
travel to work and school, and additional trip purposes. Cycling levels
were taken from the POWSCAR data, which includes the number of people
travelling by each mode between each OD pair.

@@

\begin{itemize}
\tightlist
\item
  Calculate distance decay curves for Dublin for the base year (2016,
  using POWSCAR data) by fitting a model to the relevant OD data after
  it has been converted to a route network datasest
\item
  Apply the Near Market model to the hilliness and distance values for
  each county during the build process
\item
  Add the current level of cycling to the Near Market model
\end{itemize}

@@

The Climate Action Plan scenario is loosely based on the Irish
Government's
\href{https://www.gov.ie/en/publication/6223e-climate-action-plan-2021/}{Climate
Action Plan 2021}, which contains policies for action to achieve a 51\%
reduction in overall GHG emissions by 2030, on a path towards net-zero
by 2050. For transport, this includes 500,000 extra walking, cycling and
public transport trips per day by 2030. In terms of car travel, the
target is to ``Increas{[}e{]} the proportion of kilometres driven by
passenger electric cars to between 40 and 45\% by 2030, in addition to a
reduction of 10\% in kilometres driven by the remaining internal
combustion engine cars.'' This equates to a 5.5 to 6\% reduction in
total car km driven.

@@

The main result of the CRUSE project is an open access web application
and datasets on current and potential future cycling levels in Ireland,
with evidence provided at county and route segment levels. Like the PCT
on which CRUSE builds, a key feature of the results is that they are
systematic and available for all counties, and most routes on which
cycling is permitted, in Ireland. The full set of results are therefore
too extensive to present in their entirety in this paper. Through the
interactive web application, users can explore the data to generate the
results that are most relevant to their needs, whether that is finding
the cycling potential on a particular road or finding `weak links' or
barriers in the cycle network associated with a particular school, work
place or other destination.

@@

To illustrate the results in urban areas, we took the top 4 cities in
Ireland by population: Dublin City, Cork, Limerick and Galway (with city
populations of 0.6, 0.2, 0.1 and 0.1 million respectively). The results
for central areas demarcated with circular areas with a radius of 3km
for these cities are presented in Figure~\ref{fig-city-results}.

@@

\caption{\label{fig-city-results}Fastest, balanced and quietest cycle
networks within 3km circular areas containing the four largest cities in
Ireland. The networks are coloured by cycle friendliness and with widths
varying in proportion to the potential level of cycling under the `Go
Dutch' scenario.}

@@

Aside from the widely varying sizes and route networks of each city,
with Dublin having by far the densest network, it is clear from the
visualisation of cycling friendliness that there is a lot of room for
improvement in all cities. Dublin has had received most investment in
cycling infrastructure and this is apparent from the high proportion of
the route network that is green, even in the `Fastest' route network
representing routes that prioritise speed and directness over comfort
and safety. Dublin also has the highest mode share of cycling of the
four cities, adding to the evidence base showing that investment in
cycling infrastructure is effective in increasing cycling uptake.

@@

The CRUSE Tool provides evidence on current and potential future cycling
levels across Ireland down to the street level, with potential assigned
to `fastest', `balanced' and `quietest' route networks. This provision
of multiple scenarios of behaviour changes \emph{and} multiple scenarios
of investment, for example in cycling infrastructure next to major roads
vs quiet residential streets, is a key feature of the CRUSE Tool.

@@

The CRUSE Tool can help both in terms of describing the current
situation, and also prioritise investment in those direct routes with
high cycling potential that lack adequate cycling infrastructure,
highlighted by \citet{caulfield2012}. It will also support reporting for
the RISM Directive on road casualties as part of the road safety
management system.

@@

\begin{itemize}
\tightlist
\item
  The results are \emph{open access} meaning that any stakeholder in the
  planning system can access the evidence. This will help to democratise
  the transport planning process and make wider conversations about
  transport planning more evidence-based and less polarised
  \citep{lovelace2020}, something that is particularly relevant given
  the potentially polarising nature of pro-cycling interventions
  \citep{wild2017}.
\item
  The results are fully reproducible (code to be released pending
  sign-off by TII's IT team), preventing `cloud lock in' to a
  potentially monopolistic consultancy, and encouraging input from the
  wider open source community \citep{lovelace2021, dhir2017}.
\item
  By covering a wide range of trip purposes --- not just travel to work
  \citep{lovelace2017, heinen2010} and travel school \citep{goodman2019}
  as covered in previously published research on open access national
  cycle netework planning tools --- the results capture a higher
  proportion of cycling potential including key rural trips which are
  often under-represented in models of active travel.
\end{itemize}

@@

The CRUSE Tool is an open access web application for strategic cycle
network planning and prioritisation of road safety interventions across
Ireland. Building on previous work, it provides a nationally consistent
evidence base, providing valuable insights to planners and other
stakeholders in the transport planning process, at national and local
levels. Because the results are available at the route segment level,
the tool can be used to identify `weak links' in the cycle network, and
to prioritise investment in cycling infrastructure. Furthermore, the
provision of the evidence in a free and publicly available website,
hosted at \href{https://cruse.bike}{cruse.bike}, means that it can be
used by anyone, encouraging wider participation and more evidence-based
debate about transport planning.

@@

More broadly, the tool's outputs are limited to just one more of active
travel (cycling), ignoring walking and wheeling, including wheelchair
use and a range of wheeled devices such as scooters, bike trailers and
e-cargo bikes that can be used to escort children to school. Sustainable
transport policies should plan for walking, wheeling and cycling, and
there are many co-benefits of broadly defined active travel
interventions that benefit all active modes, such as measures to reduce
heavy motor traffic speeds and volumes in areas and along corridors with
high active travel potential. This raises the question of whether other
active modes should be incorporated into the results using the OD-based
approach outlined in this paper, or whether different modelling
approaches are needed to properly capture the shorter distance trips
typically made by walking \citep{cooper2018}. Furthermore, the estimates
of cycling presented in the CRUSE Tool omit multi-modal trips including
public transport, and omit trip chaining, due to the need to capture a
large portion of cycling potential within the resource constraints of
the project.

>>>>

\begin{abstract}
Legislation mandates improved road safety and decarbonisation of
predominantly road-based transport network in Ireland and other
countries. Increasing cycling transport, making it safer and more
attractive, is essential to this end. Cycling in Ireland represents only
3\% of total modal share as of the latest Census (2016), but accounts
for 20\% of serious injuries and 7\% of all fatalities. A planning tool
based on high resolution datasets on baseline and potential cycling
levels could greatly assist in achieving this goal.

This paper describes the design, features and potential use of the Cycle
Route Acceptance and Scenario Estimation (CRUSE) Tool for Ireland was
commissioned to support cycle network planning, safety interventions,
business cases and project evaluation. The resulting web application is
available at \url{https://cruse.bike/}, enabling planners, engineers,
and other stakeholders to make more evidence-based decisions. CRUSE goes
beyond previous work in several important ways. It models networks at
high levels of geographic resolution, simulates a wide range of trips
including recreational, social, shopping and personal trips, in addition
to trips to work and primary, secondary and tertiary education
locations. The tool provides key `denominator' information to enable
reporting of collision rates per billion km cycled.

@@

\begin{keyword}
    Cycling \sep Open-source \sep Road Safety \sep 
    Active Travel
\end{keyword}
\end{frontmatter}
    \ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[breakable, enhanced, frame hidden, boxrule=0pt, interior hidden, borderline west={3pt}{0pt}{shadecolor}, sharp corners]}{\end{tcolorbox}}\fi

@@

Transport is responsible for 23\% of global greenhouse gas (GHG)
emissions, 70\% of which are from road transport. Nearly half of these
emissions (around 10\% of global emissions) are from passenger cars
\citep{jaramillo2022}. The transport system encourages, enables and in
some cases enforces unsustainable lifestyles. Services that are only
accessible by car lock-in car dependency
\citep{gray2001, shergold2012, motte-baumvol2010}.

@@

Another deadly consequence of inefficient transport systems is tackled
in the Road Infrastructure Safety Management (RISM) directive
(2008/96/EC), which requires member states to implement a road safety
management system (RSMS) for all public roads. Specifically, ``Member
States shall ensure that the ranking of high accident concentration
sections and the network safety ranking are carried out''
\citep{directiv2008}. Given that `safety' in this context is usefully
quantified as the number of people killed and seriously injured (KSI)
per distance traveled, the directive requires estimation of distance
traveled by mode, down to the road link level. For active modes, about
which there is a paucity of data compared with motorized modes, this is
a major challenge. Better data to inform road safety policies and
interventions is a motivation for the CRUSE Tool outlined in this paper,
and modeling active travel more broadly.

@@

At the regional level, the recently published Greater Dublin Area
Transport Strategy \citep{greater} reinforces these findings: ``nearly a
quarter of adults cycle at least once a week in the Dublin Metropolitan
Area'' with cycling in the Dublin area taking up to 60,000 cars off the
road today. Extrapolating this on a per population basis across Ireland,
with around 40\% of the population living in Dublin, suggests that
around 150,000 cars could be removed nationwide just by achieving Dublin
levels of cycling in all counties (notwithstanding existing cycling
trips and differences in trip distances). 7 in 10 trips in Ireland are
by car \citep{national} . Cycling has the potential to replace a large
proportion of these trips: ``A high priority must also be given to
cyclists, because trips by this mode have the potential to replace trips
by private car, most specifically for short to medium distance trips,
but increasingly for longer trips as e-bikes extend the range of this
mode'' \citep{greater}.

@@

This paper aims to describe the design, features and potential use of
the CRUSE Tool. In Section~\ref{sec-methods}, we outline the methods
used to generate the evidence presented in the CRUSE Tool. In
Section~\ref{sec-results}, we present the results of the CRUSE Tool,
including estimates of current cycling levels and future cycling
potential at the national, regional and local levels. In
Section~\ref{sec-discussion}, we discuss the implications of the results
for policy and practice. In Section~\ref{sec-conclusions}, we conclude
the paper.

@@

The methods used to generate the evidence presented in the CRUSE Tool
build on the Propensity to Cycle Tool (PCT), which was originally funded
by the UK's Department for Transport and developed by a multi-university
team. An important feature of the PCT is that it is open source and
publicly available (at \href{https://www.pct.bike/}{www.pct.bike}),
allowing its use by all stakeholders in the transport planning process
\citep{lovelace2017}. The PCT approach has had major policy impacts, as
outlined in Research Excellence Framework (REF) impact case studies,
which demonstrate that the tool ``revolutionized strategic cycle
planning in England and Wales'' by overcoming the barriers to cycling
investment imposed by lack of evidence on cycling
potential\citep{lovelace2023}.

The first version of the PCT was based on current and future potential
uptake of cycling for single stage \emph{travel to work} at desire line,
zone, route, and route network levels \citep{lovelace2016}. It was
launched in April 2017 as the government-endorsed tool for strategic
cycle network planning and part of the Cycling and Walking Investment
Strategy \citep{cycling2017}. Extensions of the PCT approach have
included estimation of benefits at the individual level
\citep{woodcock2018}, addition of travel to school network
\citep{goodman2019}, and improved modeling of impacts on health,
environmental and distributional outcomes \citep{woodcock2021}.
Initially developed just for England, the PCT was extended to cover all
of Wales (for commuter data only) in 2018.

@@

A feature of active travel interventions is that they require dense
networks of routes to be effective \citep{parkin2018}. This means that
high levels of geographic resolution are needed in the data used to
estimate cycling potential. However, datasets on travel patterns are
often only available at the level of administrative zones. The Central
Statistics Office (CSO) in Ireland provides Place of Work, School or
College Census of Anonymized Records
(\href{https://www.cso.ie/en/census/census2016reports/powscar/}{POWSCAR})
data on the number of people traveling to work and school at the
Electoral Division (ED) level, for example.

In the PCT, the method used to convert OD data to route networks was to
calculate a single route between the population weighted centroids of
the zones associated with each OD pair. This method works fine when the
OD data represents movement between small areas, but was not appropriate
for generating route networks from the POWSCAR data because zone
centroids are so far apart that the resulting route networks would be
sparse and unrealistic. To tackle this issue we developed a new method
for OD data disaggregation called `jittering' \citep{lovelace2022}. The
method works by first disaggregating the OD data based on a
`disaggregation threshold' and then assigning each disaggregated
`sub-OD' pair to `subpoints' within each zone. As shown in
Figure~\ref{fig-dublin}, the resulting route networks are dense, even in
rural areas.

@@

However, travel to school and work only constitute around only 30\% of
all trips in Ireland. To overcome this issue, we developed a spatial
interaction modeling methodology to estimate the number of trips between
each OD pair for additional trip purposes. The classification of trip
purposes used in the CRUSE Tool was guided mainly by the trip purpose
classification found within the National Household Travel Survey (NHTS),
but with the addition of categories based on the comprehensive POWSCAR
data, and the need to include recreational trips and multi-stage trips
(not yet implemented). An overview of the trip purposes used in CRUSE is
presented in Figure~\ref{fig-trip-purposes}.

@@

We calculated three network types for each disaggregated (`jittered') OD
pair: `Fastest', `Balanced' and `Quietest'. As outlined on the
CycleStreets'
\href{https://www.cyclestreets.net/help/journey/howitworks/}{website},
the `Fastest' routes minimize journey time, accounting for traffic
lights and surface type. The `Quietest' route type minimizes busy
sections of road, allowing high `diversion factors' from the fastest
route on quieter but less direct ways (often avoiding roads and
interactions with motor vehicles altogether where possible). The
`Balanced' route type is a compromize between the two, minimising
journey time while avoiding the busiest roads. Allowing users to switch
between these three route types enables them to consider the trade-offs
between directness and `cycle friendliness' when planning new
infrastructure, as shown in Figure~\ref{fig-route-types}.

@@

The datasets outlined above were used to generate estimates of cycling
currently (the `Baseline' scenario) and under four scenarios of cycling
potential: Near Market, Climate Action Plan, Go Dutch, and Ebike. Each
of these scenarios is described below, as illustrated in
Figure~\ref{fig-scenarios}.

@@

The Baseline scenario approximates current cycling levels. As outlined
in Section~\ref{sec-trip-purposes}, the Baseline scenario includes
travel to work and school, and additional trip purposes. Cycling levels
were taken from the POWSCAR data, which includes the number of people
traveling by each mode between each OD pair.

@@

\begin{itemize}
\tightlist
\item
  Calculate distance decay curves for Dublin for the base year (2016,
  using POWSCAR data) by fitting a model to the relevant OD data after
  it has been converted to a route network dataset
\item
  Apply the Near Market model to the hilliness and distance values for
  each county during the build process
\item
  Add the current level of cycling to the Near Market model
\end{itemize}

@@

The Climate Action Plan scenario is loosely based on the Irish
Government's
\href{https://www.gov.ie/en/publication/6223e-climate-action-plan-2021/}{Climate
Action Plan 2021}, which contains policies for action to achieve a 51\%
reduction in overall GHG emissions by 2030, on a path towards net-zero
by 2050. For transport, this includes 500,000 extra walking, cycling and
public transport trips per day by 2030. In terms of car travel, the
target is to ``Increas{[}e{]} the proportion of kilometers driven by
passenger electric cars to between 40 and 45\% by 2030, in addition to a
reduction of 10\% in kilometers driven by the remaining internal
combustion engine cars.'' This equates to a 5.5 to 6\% reduction in
total car km driven.

@@

The main result of the work presented in this paper is an open access
web application and datasets on current and potential future cycling
levels in Ireland, with evidence provided at county and route segment
levels. Like the PCT on which CRUSE builds, a key feature of the results
is that they are systematic and available for all counties, and most
routes on which cycling is permitted, in Ireland. The full set of
results are therefore too extensive to present in their entirety in this
paper. Through the interactive web application, users can explore the
data to generate the results that are most relevant to their needs,
whether that is finding the cycling potential on a particular road or
finding `weak links' or barriers in the cycle network associated with a
particular school, work place or other destination.

@@

To illustrate the results in urban areas, we took the top 4 cities in
Ireland by population: Dublin City, Cork, Limerick and Galway (with city
populations of 0.6, 0.2, 0.1 and 0.1 million respectively). In
Figure~\ref{fig-city-results} we present the cycle route networks within
a 3km radius circle centred on each of these cities.

@@

\caption{\label{fig-city-results}Fastest, balanced and quietest cycle
networks within 3km circular areas containing the four largest cities in
Ireland. The networks are coloured by cycle friendliness, with greener
segments representing more cycle friendly segments. Widths vary in
proportion to the potential level of cycling under the `Go Dutch'
scenario.}

@@

Aside from the widely varying sizes and route networks of each city,
with Dublin having by far the densest network, it is clear from the
visualisation of cycling friendliness that there are many gaps in the
networks. Dublin has had received most investment in cycling
infrastructure and this is apparent from the high proportion of the
route network that is green, even in the `Fastest' route network
representing routes that prioritize speed and directness over comfort
and safety. Dublin also has the highest mode share of cycling of the
four cities, adding to the evidence base showing that investment in
cycling infrastructure is effective in increasing cycling uptake.

@@

The aim of the paper was to describe the design, features and potential
use of the CRUSE Tool. As outlined in the previous section, the CRUSE
Tool provides evidence on current and potential future cycling levels
across Ireland down to the street level, with potential assigned to
`fastest', `balanced' and `quietest' route networks. This provision of
multiple scenarios of behaviour change \emph{and} multiple scenarios of
investment, for example in cycling infrastructure next to major roads vs
quiet residential streets, is a key feature of the CRUSE Tool.

@@

The CRUSE Tool can help both in terms of describing the current
situation, and also prioritise investment in those direct routes with
high cycling potential that lack adequate cycling infrastructure
\citep{caulfield2012}. It will also support reporting for the RISM
Directive on road casualties as part of the road safety management
system.

@@

\begin{itemize}
\tightlist
\item
  The results are \emph{open access} meaning that any stakeholder in the
  planning system can access the evidence. This will help to democratize
  the transport planning process and make wider conversations about
  transport planning more evidence-based and less polarized
  \citep{lovelace2020}, something that is particularly relevant given
  the potentially polarizing nature of pro-cycling interventions
  \citep{wild2017}.
\item
  The results are fully reproducible (code to be released pending
  sign-off by TII's IT team), preventing `cloud lock in' to a
  potentially monopolistic consultancy, and encouraging input from the
  wider open source community \citep{lovelace2021, dhir2017}.
\item
  By covering a wide range of trip purposes --- not just travel to work
  \citep{lovelace2017, heinen2010} and travel school \citep{goodman2019}
  as covered in previously published research on open access national
  cycle network planning tools --- the results capture a higher
  proportion of cycling potential including key rural trips which are
  often under-represented in models of active travel.
\end{itemize}

@@

The CRUSE Tool is an open access web application for strategic cycle
network planning and prioritization of road safety interventions across
Ireland. Building on previous work, it provides a nationally consistent
evidence base, providing valuable insights to planners and other
stakeholders in the transport planning process, at national and local
levels. Because the results are available at the route segment level,
the tool can be used to identify `weak links' in the cycle network, and
to prioritize investment in cycling infrastructure. Furthermore, the
provision of the evidence in a free and publicly available website,
hosted at \href{https://cruse.bike}{cruse.bike}, means that it can be
used by anyone, encouraging wider participation and more evidence-based
debate about transport planning.

@@

More broadly, the tool's outputs are limited to just one mode of active
travel (cycling), ignoring walking and wheeling, including wheelchair
use and a range of wheeled devices such as scooters, bike trailers and
e-cargo bikes that can be used to escort children to school. Sustainable
transport policies should plan for walking, wheeling and cycling, and
there are many co-benefits of broadly defined active travel
interventions that benefit all active modes, such as measures to reduce
heavy motor traffic speeds and volumes in areas and along corridors with
high active travel potential. This raises the question of whether other
active modes should be incorporated into the results using the OD-based
approach outlined in this paper, or whether different modeling
approaches are needed to properly capture the shorter distance trips
typically made by walking \citep{cooper2018}. Furthermore, the estimates
of cycling presented in the CRUSE Tool omit multi-modal trips including
public transport, and omit trip chaining, due to the need to capture a
large portion of cycling potential within the resource constraints of
the project.
